---
title: "A general pipeline for looking at mutli-trait multiple-qtl models"
author: "JT Lovell"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width = 7, fig.height = 7, warning = FALSE)
library(knitr)
library(qtl)
library(qtlTools)
library(plyr)
```

```{r env.set.up, eval = FALSE}
library(devtools)
install_github("jtlovell/qtlTools")
```

```{r env.set.up2}
library(qtlTools)
library(qtl)
library(ggplot2)
library(lsmeans)
library(qtlpvl)
```

### Simulate a few phenotypes in a five-chromosome map
For this tutorial, we will use the dataset `multitrait` from R/qtl. 
```{r}
data(multitrait)
cross<-multitrait
cross<-calc.genoprob(cross)
phes<-phenames(cross)
```

In R/qtl there is a pipeline to analyze [multiple traits in a single model](http://www.rqtl.org/tutorials/MQM-tour.pdf). Here, we will discuss analyzing them independently, then concatenating the results

### Scanones
In the simplest case, we can run a bunch of scanones
```{r}
s1<-scanone(cross, pheno.col = phes, method = "hk")
cols<-rainbow(length(phes))
plot(s1, type="n", ylim = c(0,max(as.matrix(s1[,-c(1:2)]))), 
     ylab = "LOD Score", main = "all phenotypes plotted together")
for(i in 1:length(phes)) plot(s1, add=T, lodcolumn = i, col = cols[i])
```

### QTL peaks from scanones
This simplest way to look at these data is to pull out only the significant QTL peaks 
```{r}
perms<-scanone(cross, pheno.col = phes, method = "hk", n.perm = 100, verbose=F)
print(pullSigQTL(cross, pheno.col=phes,
                 s1.output = s1,  perm.output = perms, 
                 returnQTLModel=FALSE, alpha = 0.05,
                 controlAcrossCol = TRUE))
```

However, if we want to look at model statistics etc. it is best to create multiple qtl models for each phenotype
```{r}
mods<-pullSigQTL(cross, pheno.col=phes,
                 s1.output = s1,  perm.output = perms, 
                 returnQTLModel=TRUE, alpha = 0.05,
                 controlAcrossCol = TRUE)
```

This returns a list of fitted QTL models, which we could use to calculate confidence intervals etc. 
```{r}
print(mods[1:5])
```

### Multiple QTL models
The approach above is nice because it is super fast, but it does not necessarily produce the best models (e.g. there can only be one QTL/chromosome). One could use stepwiseqtl to generate the models instead, however, scantwo permutations are needed for this appraoch, which is slow for many traits. 

For the purposes of this tutorial, we will use the models generated through scanones to make statistical inference. 
The function `qtlStats` takes a qtl model and returns fitted statistics. 
For a single phenotype, we'd run it as follows:
```{r, warning = FALSE}
mod<-mods[["X3.Methylsulfinylpropyl"]]
form<-paste("y ~",paste(mod$altname, collapse = " + "))
mod<-refineqtl(cross, qtl = mod, formula = form, verbose=F, method="hk")
qtlStats(cross, 
         pheno.col = "X3.Methylsulfinylpropyl", 
         form = form, 
         mod = mod)
```

Since mods is a list, we can loop through it using lapply
First toss out an NULL QTL models
```{r}
mods<-mods[sapply(mods,length)!=1] # a NULL QTL model has length = 1, toss these
```

Next do the loop, where `x` represents each phenotype. 
```{r, warning = FALSE}
stats<-lapply(names(mods), function(x){
  mod<-mods[[x]]
  form<-paste("y ~",paste(mod$altname, collapse = " + "))
  mod<-refineqtl(cross, qtl = mod, formula = form, 
                verbose=F, method="hk", pheno.col = x)
  qtlStats(cross, 
         pheno.col = x, 
         form = form, 
         mod = mod)
})
stats<-do.call(rbind, stats)
```

### Plotting multiple QTL models
The most basic way to present these data is to show the position of QTL intervals on a genetic map.
`qtlTools::segmentsOnMap` provides one method to do this.
```{r}
segmentsOnMap(cross=cross, phe=stats$phenotype, 
              chr=stats$chr, 
              l = stats$lowposition, 
              h =stats$highposition, segSpread = .05,legendCex=.5,
              lwd = 2, legendPosition = "bottom", leg.inset=.01)
```
