---
title: "Comparison of permutation methods"
author: "JT Lovell"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width = 7, fig.height = 7, warning = FALSE)
library(knitr)
library(qtl)
library(qtlTools)
```

```{r env.set.up}
library(devtools)
install_github("jtlovell/qtlTools")
```

```{r env.set.up2}
library(qtlTools)
library(qtl)
```

### Get the basic scripts from Karl Broman's github
```{r get source}
source("https://raw.githubusercontent.com/kbroman/qtl/master/R/util.R")
source("https://raw.githubusercontent.com/kbroman/qtl/master/R/calc.pairprob.R")
source("https://raw.githubusercontent.com/kbroman/qtl/master/R/xchr.R")
```


### Load the simulated F2 data 
```{r load data}
data(fake.f2)
covar<-data.frame(covar = fake.f2$phe$sex)
fake.f2<-calc.genoprob(fake.f2)
```

### scanone permutation comparison
#### normal permutations
```{r scanone normal}
set.seed(42)
perms0<-scanone(fake.f2, pheno.col="phenotype", addcovar=covar,
                intcovar=covar, perm.strata=covar[,1],
                n.perm=100, verbose = F)
summary(perms0)
```

#### GWERk scanone permutations
```{r scanone gwerk}
set.seed(42)
perms1<-scanone.GWERk(fake.f2, pheno.col="phenotype", 
                      addcovar=covar, intcovar=covar, 
                      perm.strata=covar[,1],
                      n.perm=100, GWERk=1, verbose = F)
summary(perms1)

plot(as.numeric(perms0), as.numeric(perms1), 
     xlab="standard perms", ylab = "GWER perms", 
     main = "scanone permutation comparison")
abline(a=0,b=1, lty=3)
```


### scantwo permutation comparison 
Here, we will just use the 1st and 13th chromosomes, to 
speed up computational time.
#### normal permutations
```{r scantwo normal}
set.seed(42)
perms2.0<-scantwo(fake.f2, chr = c(1,13), pheno.col="phenotype", 
                  addcovar=covar, intcovar=covar, 
                  perm.strata=covar[,1],
                  n.perm=50, verbose = F)
summary(perms2.0)
```

#### GWER k=1 permutations
```{r scantwo gwerk}
set.seed(42)
perms2.1<-scantwo.GWERk1(fake.f2, chr = c(1,13), 
                         pheno.col="phenotype", 
                         addcovar=covar, intcovar=covar, 
                         perm.strata = covar[,1],
                         n.perm=50, verbose = F)
summary(perms2.1)

plot(as.numeric(perms2.0$full), as.numeric(perms2.1$full), 
     xlab="standard perms", ylab = "GWER perms", 
     main = "scantwo permutation comparison")
abline(a=0,b=1, lty=3)
```

### stepwiseqtl comparison
#### Normal stepwiseqtl
```{r stepwise}
pens = calc.penalties(perms2.1, alpha = 0.2)

outsw <- stepwiseqtl(fake.f2, pheno.col="phenotype", chr = c(1,13),
                     max.qtl=3, method="hk", keeptrace=TRUE, covar=covar,
                     penalties = pens, verbose = F)
summary(outsw)
summary(fitqtl(fake.f2, pheno.col="phenotype", covar=covar,
               formula=formula(outsw), qtl = outsw, 
               method="hk"))

```

#### stepwiseqtl forcing covariate interaction
```{r}
outsw.int <- stepwiseqtl.int(fake.f2, pheno.col="phenotype", chr = c(1,13),
                             max.qtl=3, method="hk", keeptrace=TRUE, 
                             covar=covar, verbose = F)
summary(outsw.int)
summary(fitqtl(fake.f2, pheno.col="phenotype", 
               formula=formula(outsw.int), qtl = outsw.int, 
               covar = covar, method="hk"))

par(mfrow=c(2,1))
plotLodProfile(outsw, main = "normal stepwise with additive covariate")
plotLodProfile(outsw.int, main = "normal stepwise with interactive covariate")
```

